### 2014-04-29 doc/coppermtn_multilevel/tech_report/multilevel_mc.tex
adding initial coarse grid results to the mlmc tech report

\subsection{Heterogeneous Results}
We next solve the same linear problem as before using $M=2$ but this
time with a constant forcing of $\ve{b}=1.0$ such that:
\begin{equation}
  \ve{A}\ve{x} = \ve{b}\:.
  \label{eq:heterogeneous_problem}
\end{equation}
We will choose an initial guess of $\ve{0}$ for these calculations as
the solution is non-trivial and fix $N$ at 10,000 while increasing the
number of levels. Table~\ref{tab:heterogeneous_fixed_t} gives the
results of these calculations. Here the relative error error
calculations were computed using an exact solution computed using a
minimal residual iteration converged to a residual norm tolerance of
$\sn{1}{-12}$.
\begin{table}[h!]
  \begin{center}
    \begin{tabular}{cllllllll}\hline\hline
      \multicolumn{1}{c}{\textbf{Levels}} &
      \multicolumn{1}{l}{\textbf{Samples}} &
      \multicolumn{1}{l}{\textbf{$||\ve{r}||_{\infty}$}} &
      \multicolumn{1}{l}{\textbf{$||\ve{r}||_{2}$}} &
      \multicolumn{1}{l}{\textbf{$||\ve{e}||_{\infty}$}} &
      \multicolumn{1}{l}{\textbf{$||\ve{e}||_{2}$}} &
      \multicolumn{1}{l}{\textbf{Time (s)}} &
      \multicolumn{1}{l}{\textbf{RRESFOM}} &
      \multicolumn{1}{l}{\textbf{RERRFOM}} \\
      \hline
      1 & 10,000 & 0.025 & 0.223 & 0.241 & 2.97 & 6.77 & 1.0 & 1.0 \\
      2 & 17,596 & 0.044 & 0.339 & 0.277 & 5.49 & 7.19 & 0.31 & 0.71 \\
      3 & 29,571 & 0.050 & 0.365 & 0.131 & 1.66 & 7.21 & 0.23 & 3.18 \\
      4 & 53,295 & 0.062 & 0.483 & 0.081 & 0.96 & 7.29 & 0.16 & 8.28 \\
      5 & 92,301 & 0.086 & 0.662 & 0.105 & 0.88 & 7.13 & 0.08 & 5.00 \\
      6 & 185,266 & 0.120 & 0.891 & 0.102 & 0.72 & 7.10 & 0.04 & 5.32 \\
      7 & 355,543 & 0.069 & 0.513 & 0.056 & 0.51 & 6.92 & 0.13 & 18.12 \\
      8 & 742,337 & 0.093 & 0.654 & 0.081 & 0.67 & 7.07 & 0.07 & 8.47 \\
      9 & 1,624,121 & 0.122 & 0.758 & 0.107 & 0.83 & 7.08 & 0.04 &
      4.85 \\
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Geometric Multilevel Monte Carlo results for
      $\ve{b} = 1$ with $M = 2$, $t \approx 7s$ and run times reported
      in seconds.} \textit{RRESSFOM values are the relative figure of
      merit computed using the infinity norm of the residual. RERRFORM
    values are the relative figure of merit computed using the
    infinity norm of the absolute error vector.}}
  \label{tab:heterogeneous_fixed_t}
\end{table}
Based on the residual norm values, it would seem as though the 1 level
result was superior as better convergence was achieved in the same
amount of compute time using residual norm metrics. However, the
structure of the actual solutions would indicate
otherwise. Figure~\ref{fig:heterogeneous_ml_solutions} gives the
results for with 1-9 levels all plotted against the exact
solution. Clearly, the estimate produced by the 9 level computation is
more representative of the exact solution than the single level scheme
as indicated by the error norms due to the well-converged results
provided coarse level estimates. Furthermore, the error norms also
indicate a more efficient solution scheme when multiple levels are
used showing similar relative figure of merit metrics for the error
norm computations.

\begin{figure}[h!]
  \centering
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{7_sec_1_level.pdf}
    \caption{\textbf{1 level.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{7_sec_2_level.pdf}
    \caption{\textbf{2 levels.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{7_sec_3_level.pdf}
    \caption{\textbf{3 levels.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{7_sec_4_level.pdf}
    \caption{\textbf{4 levels.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{7_sec_5_level.pdf}
    \caption{\textbf{5 levels.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{7_sec_6_level.pdf}
    \caption{\textbf{6 levels.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{7_sec_7_level.pdf}
    \caption{\textbf{7 levels.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{7_sec_8_level.pdf}
    \caption{\textbf{8 levels.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{7_sec_9_level.pdf}
    \caption{\textbf{9 levels.}}
  \end{subfigure}
  \caption{\textbf{Solution of the Poisson problem using the geometric
      multilevel Monte Carlo method with 1-9 levels and approximately
      7 seconds of compute time.} \textit{Increasing the number of
      levels in the computation improves convergence with respect to
      the relative solution error but convergence with respect to the
      linear problem residual is worse.}}
  \label{fig:heterogenous_ml_solutions}
\end{figure}

@@

%%---------------------------------------------------------------------------%%
\section{Coarse Grid Monte Carlo Acceleration}
Based on the above results we then consider using only the coarse grid
calculations as the acceleration. Consider the following results where
a coarse grid solution has been produced and then simply prolongated
back to the fine grid rather than solving all levels as a simple
coarse approximation of the solution.

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{cllllllll}\hline\hline
      \multicolumn{1}{c}{\textbf{Levels}} &
      \multicolumn{1}{l}{\textbf{Samples}} &
      \multicolumn{1}{l}{\textbf{$||\ve{r}||_{\infty}$}} &
      \multicolumn{1}{l}{\textbf{$||\ve{r}||_{2}$}} &
      \multicolumn{1}{l}{\textbf{$||\ve{e}||_{\infty}$}} &
      \multicolumn{1}{l}{\textbf{$||\ve{e}||_{2}$}} &
      \multicolumn{1}{l}{\textbf{Time (s)}} &
      \multicolumn{1}{l}{\textbf{RRESFOM}} &
      \multicolumn{1}{l}{\textbf{RERRFOM}} \\
      \hline
      1 & 10,000 & 0.025 & 0.223 & 0.241 & 2.97 & 6.77 & 1.0 & 1.0 \\
      2 & 17,596 & 0.044 & 0.339 & 0.277 & 5.49 & 7.19 & 0.31 & 0.71 \\
      3 & 29,571 & 0.050 & 0.365 & 0.131 & 1.66 & 7.21 & 0.23 & 3.18 \\
      4 & 53,295 & 0.062 & 0.483 & 0.081 & 0.96 & 7.29 & 0.16 & 8.28 \\
      5 & 92,301 & 0.086 & 0.662 & 0.105 & 0.88 & 7.13 & 0.08 & 5.00 \\
      6 & 185,266 & 0.120 & 0.891 & 0.102 & 0.72 & 7.10 & 0.04 & 5.32 \\
      7 & 355,543 & 0.069 & 0.513 & 0.056 & 0.51 & 6.92 & 0.13 & 18.12 \\
      8 & 742,337 & 0.093 & 0.654 & 0.081 & 0.67 & 7.07 & 0.07 & 8.47 \\
      9 & 1,624,121 & 0.122 & 0.758 & 0.107 & 0.83 & 7.08 & 0.04 &
      4.85 \\
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Coarse grid results for $\ve{b} = 1$ with $M = 2$,
      $N = 10,000$ and run times reported in seconds.}
    \textit{RRESSFOM values are the relative figure of merit computed
      using the infinity norm of the residual. RERRFORM values are the
      relative figure of merit computed using the infinity norm of the
      absolute error vector.}}
  \label{tab:heterogeneous_coarse_grids}
\end{table}

\begin{figure}[h!]
  \centering
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{7_sec_1_level.pdf}
    \caption{\textbf{Level 1 solution.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{cg_2_level.pdf}
    \caption{\textbf{Level 2 solution.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{cg_3_level.pdf}
    \caption{\textbf{Level 3 solution.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{cg_4_level.pdf}
    \caption{\textbf{Level 4 solution.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{cg_5_level.pdf}
    \caption{\textbf{Level 5 solution.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{cg_6_level.pdf}
    \caption{\textbf{Level 6 solution.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{cg_7_level.pdf}
    \caption{\textbf{Level 7 solution.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{cg_8_level.pdf}
    \caption{\textbf{Level 8 solution.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{cg_9_level.pdf}
    \caption{\textbf{Level 9 solution.}}
  \end{subfigure}
  \caption{\textbf{Solution of the Poisson problem using the geometric
      multilevel Monte Carlo method with 1-9 levels and approximately
      7 seconds of compute time.} \textit{Increasing the number of
      levels in the computation improves convergence with respect to
      the relative solution error but convergence with respect to the
      linear problem residual is worse.}}
  \label{fig:heterogenous_ml_solutions}
\end{figure}

>>>>

\subsection{Heterogeneous Results}
We next solve the same linear problem as before using $M=2$ but this
time with a constant forcing of $\ve{b}=1.0$ such that:
\begin{equation}
  \ve{A}\ve{x} = \ve{b}\:.
  \label{eq:heterogeneous_problem}
\end{equation}
We will choose an initial guess of $\ve{0}$ for these calculations as
the solution is non-trivial and fix $N$ at 10,000 while increasing the
number of levels. Table~\ref{tab:heterogeneous_fixed_t} gives the
results of these calculations. Here the relative error error
calculations were computed using an exact solution computed using a
minimal residual iteration converged to a residual norm tolerance of
$\sn{1}{-12}$.
\begin{table}[h!]
  \begin{center}
    \begin{tabular}{cllllllll}\hline\hline
      \multicolumn{1}{c}{\textbf{Levels}} &
      \multicolumn{1}{l}{\textbf{Samples}} &
      \multicolumn{1}{l}{\textbf{$||\ve{r}||_{\infty}$}} &
      \multicolumn{1}{l}{\textbf{$||\ve{r}||_{2}$}} &
      \multicolumn{1}{l}{\textbf{$||\ve{e}||_{\infty}$}} &
      \multicolumn{1}{l}{\textbf{$||\ve{e}||_{2}$}} &
      \multicolumn{1}{l}{\textbf{Time (s)}} &
      \multicolumn{1}{l}{\textbf{RRESFOM}} &
      \multicolumn{1}{l}{\textbf{RERRFOM}} \\
      \hline
      1 & 10,000 & 0.025 & 0.223 & 0.241 & 2.97 & 6.77 & 1.0 & 1.0 \\
      2 & 17,596 & 0.044 & 0.339 & 0.277 & 5.49 & 7.19 & 0.31 & 0.71 \\
      3 & 29,571 & 0.050 & 0.365 & 0.131 & 1.66 & 7.21 & 0.23 & 3.18 \\
      4 & 53,295 & 0.062 & 0.483 & 0.081 & 0.96 & 7.29 & 0.16 & 8.28 \\
      5 & 92,301 & 0.086 & 0.662 & 0.105 & 0.88 & 7.13 & 0.08 & 5.00 \\
      6 & 185,266 & 0.120 & 0.891 & 0.102 & 0.72 & 7.10 & 0.04 & 5.32 \\
      7 & 355,543 & 0.069 & 0.513 & 0.056 & 0.51 & 6.92 & 0.13 & 18.12 \\
      8 & 742,337 & 0.093 & 0.654 & 0.081 & 0.67 & 7.07 & 0.07 & 8.47 \\
      9 & 1,624,121 & 0.122 & 0.758 & 0.107 & 0.83 & 7.08 & 0.04 &
      4.85 \\
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Coarse grid results for $\ve{b} = 1$ with $M = 2$,
      $t \approx 7s$ and run times reported in seconds.}
    \textit{RRESSFOM values are the relative figure of merit computed
      using the infinity norm of the residual. RERRFORM values are the
      relative figure of merit computed using the infinity norm of the
      absolute error vector.}}
  \label{tab:heterogeneous_coarse_grids}
\end{table}

Based on the residual norm values, it would seem as though the 1 level
result was superior as better convergence was achieved in the same
amount of compute time using residual norm metrics. However, the
structure of the actual solutions would indicate
otherwise. Figure~\ref{fig:heterogeneous_ml_solutions} gives the
results for with 1-9 levels all plotted against the exact
solution. Clearly, the estimate produced by the 9 level computation is
more representative of the exact solution than the single level scheme
as indicated by the error norms due to the well-converged results
provided coarse level estimates. Furthermore, the error norms also
indicate a more efficient solution scheme when multiple levels are
used showing similar relative figure of merit metrics for the error
norm computations.

\begin{figure}[h!]
  \centering
  \begin{subfigure}[b]{0.32\textwidth}
    \includegraphics[width=\textwidth]{7_sec_1_level.pdf}
    \caption{\textbf{1 level.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\textwidth}
    \includegraphics[width=\textwidth]{7_sec_2_level.pdf}
    \caption{\textbf{2 levels.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\textwidth}
    \includegraphics[width=\textwidth]{7_sec_3_level.pdf}
    \caption{\textbf{3 levels.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\textwidth}
    \includegraphics[width=\textwidth]{7_sec_4_level.pdf}
    \caption{\textbf{4 levels.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\textwidth}
    \includegraphics[width=\textwidth]{7_sec_5_level.pdf}
    \caption{\textbf{5 levels.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\textwidth}
    \includegraphics[width=\textwidth]{7_sec_6_level.pdf}
    \caption{\textbf{6 levels.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\textwidth}
    \includegraphics[width=\textwidth]{7_sec_7_level.pdf}
    \caption{\textbf{7 levels.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\textwidth}
    \includegraphics[width=\textwidth]{7_sec_8_level.pdf}
    \caption{\textbf{8 levels.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\textwidth}
    \includegraphics[width=\textwidth]{7_sec_9_level.pdf}
    \caption{\textbf{9 levels.}}
  \end{subfigure}
  \caption{\textbf{Solution of the Poisson problem using the geometric
      multilevel Monte Carlo method with 1-9 levels and approximately
      7 seconds of compute time.} \textit{Increasing the number of
      levels in the computation improves convergence with respect to
      the relative solution error but convergence with respect to the
      linear problem residual is worse.}}
  \label{fig:heterogenous_fixed_t}
\end{figure}

@@

%%---------------------------------------------------------------------------%%
\section{Coarse Grid Monte Carlo Acceleration}
Based on the above results we then consider using only the coarse grid
calculations as the acceleration. Consider the following results given
in Table~\ref{tab:heterogeneous_cg_solutions} where a coarse grid
solution has been produced at a fixed $N=10,000$ and then simply
prolongated back to the fine grid rather than solving all levels using
Algorithm~\ref{alg:mlamc}. Figure~\ref{fig:hetergeneous_cg_solutions}
plots the coarse grid solutions prolongated back to the fine grid
along with the exact solution. We see that as we coarsen the grid
while fixing $N$, the coarse grid statistics are better than the fine
grid as there is a larger ratio of samples to grid points. The result
is a very precise coarse grid result that when prolongated to the fine
grid simply incurs the grid error associated with the linear
interpolation.

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{cllllllll}\hline\hline
      \multicolumn{1}{c}{\textbf{Levels}} &
      \multicolumn{1}{l}{\textbf{Samples}} &
      \multicolumn{1}{l}{\textbf{$||\ve{r}||_{\infty}$}} &
      \multicolumn{1}{l}{\textbf{$||\ve{r}||_{2}$}} &
      \multicolumn{1}{l}{\textbf{$||\ve{e}||_{\infty}$}} &
      \multicolumn{1}{l}{\textbf{$||\ve{e}||_{2}$}} &
      \multicolumn{1}{l}{\textbf{Time (s)}} &
      \multicolumn{1}{l}{\textbf{RRESFOM}} &
      \multicolumn{1}{l}{\textbf{RERRFOM}} \\
      \hline
      1 & 10,000 & 0.025 & 0.223 & 0.241 & 2.97 & 6.77 & 1.0 & 1.0 \\
      2 & 10,000 & 0.012 & 0.067 & 0.303 & 6.05 & 3.29 & 9.70 & 1.3 \\
      3 & 10,000 & 0.005 & 0.028 & 0.133 & 1.70 & 1.62 & 103.6 & 13.7 \\
      4 & 10,000 & 0.004 & 0.013 & 0.058 & 0.61 & 0.81 & 292.7 & 146.3 \\
      5 & 10,000 & 0.023 & 0.031 & 0.141 & 3.06 & 0.40 & 20.8 & 49.4 \\
      6 & 10,000 & 0.006 & 0.008 & 0.051 & 0.77 & 0.20 & 557.0 & 741.1 \\
      7 & 10,000 & 0.006 & 0.006 & 0.049 & 0.77 & 0.10 & 1,103 & 1,589 \\
      8 & 10,000 & 0.009 & 0.009 & 0.022 & 0.29 & 0.05 & 952.8 & 15,329 \\
      9 & 10,000 & 0.013 & 0.014 & 0.032 & 0.51 & 0.03 & 997.9 &
      14,826 \\
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Geometric Multilevel Monte Carlo results for
      $\ve{b} = 1$ with $M = 2$, $N = 10,000$ and run times reported
      in seconds.} \textit{RRESSFOM values are the relative figure of
      merit computed using the infinity norm of the residual. RERRFORM
    values are the relative figure of merit computed using the
    infinity norm of the absolute error vector.}}
  \label{tab:heterogeneous_cg_solutions}
\end{table}

\begin{figure}[h!]
  \centering
  \begin{subfigure}[b]{0.32\textwidth}
    \includegraphics[width=\textwidth]{7_sec_1_level.pdf}
    \caption{\textbf{Level 1 solution.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\textwidth}
    \includegraphics[width=\textwidth]{cg_2_level.pdf}
    \caption{\textbf{Level 2 solution.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\textwidth}
    \includegraphics[width=\textwidth]{cg_3_level.pdf}
    \caption{\textbf{Level 3 solution.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\textwidth}
    \includegraphics[width=\textwidth]{cg_4_level.pdf}
    \caption{\textbf{Level 4 solution.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\textwidth}
    \includegraphics[width=\textwidth]{cg_5_level.pdf}
    \caption{\textbf{Level 5 solution.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\textwidth}
    \includegraphics[width=\textwidth]{cg_6_level.pdf}
    \caption{\textbf{Level 6 solution.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\textwidth}
    \includegraphics[width=\textwidth]{cg_7_level.pdf}
    \caption{\textbf{Level 7 solution.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\textwidth}
    \includegraphics[width=\textwidth]{cg_8_level.pdf}
    \caption{\textbf{Level 8 solution.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.32\textwidth}
    \includegraphics[width=\textwidth]{cg_9_level.pdf}
    \caption{\textbf{Level 9 solution.}}
  \end{subfigure}
  \caption{\textbf{Solution of the Poisson problem using the geometric
      multilevel Monte Carlo method with 1-9 levels and $N = 10,000$.}
    \textit{Coarse solutions prolongated to the fine grid give better
      results in less time than fine grid solutions.}}
  \label{fig:heterogenous_cg_solutions}
\end{figure}
